environmental_data_preperation

### 2018-09-19 src/environmental_data_preperation.Rmd
winter calculation correction

# Package dependencies
You can load them using the following code which uses a function called [ipak](https://gist.github.com/stevenworthington/3178163). 
Note this function checks to see if the packages are installed first.
```{r pre-install packages, message=FALSE}
packages <- c("raster", "rgdal") 
source("../src/ipak.R")
ipak(packages)
```

@@

Spring:	1 March - 31 May (col 4:6) Summer: 1 June - 31 August (col 7:9) Autumn: 1 Sept - 31 Nov (col 10:12) Winter: 1 Dec - 28 Feb (col 13, 1:2)
```{r}
amophases <- amoraw 
amophases$YearAvg <- rowMeans(amophases[2:13], na.rm = TRUE) #the month data is in col 2-13
amophases$SpringAvg <-  rowMeans(amophases[4:6], na.rm = TRUE)
amophases$SummerAvg <-  rowMeans(amophases[7:9], na.rm = TRUE)
amophases$AutumnAvg <-  rowMeans(amophases[10:12], na.rm = TRUE)
# winter has a different format - the columns are in 1, 2 & 3 which means the rowmeans function needs a bit of tweaking
amophases$WinterAvg <- rowMeans(amophases[, c(13, 2, 3)], na.rm = TRUE) #selects just cols 12, 2, & 3
```

>>>>

# Package dependencies
You can load them using the following code which uses a function called [ipak](https://gist.github.com/stevenworthington/3178163). 
Note this function checks to see if the packages are installed first.
```{r pre-install packages, message=FALSE}
packages <- c("ncdf4", "raster", "rgdal", "useful") 
source("../src/ipak.R")
ipak(packages)
```

@@

Spring:	1 March - 31 May (col 4:6) Summer: 1 June - 31 August (col 7:9) Autumn: 1 Sept - 31 Nov (col 10:12) Winter: 1 Dec - 28 Feb (col 13 from prev year, 1:2)
```{r}
amophases <- amoraw 
amophases$YearAvg <- rowMeans(amophases[2:13], na.rm = TRUE) #the month data is in col 2-13
amophases$SpringAvg <-  rowMeans(amophases[4:6], na.rm = TRUE)
amophases$SummerAvg <-  rowMeans(amophases[7:9], na.rm = TRUE)
amophases$AutumnAvg <-  rowMeans(amophases[10:12], na.rm = TRUE)
```
winter has a different format - the columns are in 1 & 2, & 13 from the previous column, which means the rowmeans function needs a bit of tweaking...

1) shift the dec column you created down by 1
```{r}
library(useful)
amophases <- shift.column(data = amophases, columns = "Dec", len = 1, up = FALSE)
amophases
write.csv(amophases, file = "../output/env/amo.csv", row.names = FALSE) #for some reason you need to write to a csv and then reload for the next step to work properly...
```

the shifted data is in Dec.Shifted...
```{r}
amophases <- read.csv("../output/env/amo.csv")
amophases$WinterAvg <- rowMeans(amophases[, c("Dec.Shifted", "Jan", "Feb")], na.rm = TRUE) 
amophases
```
